---
title: "Parallel Computing Practical"
format: html
editor: visual
---

```{r}
#| warning: false

library(MASS)
library(doParallel)
library(foreach)
```

# Question 1

```{r}

q1 <- foreach(i = 1:100, .combine = rbind) %do% {
  samp <- rexp(100, 1)
  x_bar <- mean(samp)
  var_x <- var(samp)
  output <- c(mean = x_bar, variance = var_x)
}

q1

```

# Question 2

## Parallel processing

### Single bootstrap sample

```{r}

library(doParallel)

gal <- galaxies
n <- length(gal)

c1 <- makeCluster(2)
registerDoParallel(c1)

system.time( {
  result_parallel <- foreach(i = 1:1000, .combine = c) %dopar% {
  samp <- sample(gal, n, replace = TRUE)
  med <- median(samp)
  }
})

```

```{r}
stopCluster(c1)
```

### 1000 bootstrap samples

```{r}

c2 <- makeCluster(2)
registerDoParallel(c2)

system.time( {
  results_large_p <- foreach(i = 1:100, .combine = c) %dopar% {
  vec <- numeric(1000)
  
  for (j in 1:1000) {
    s <- sample(gal, n, replace = TRUE)
    vec[j] <- median(s)
  }
  
  }
})

```

```{r}
stopCluster(c2)
```

## Serial processing

### Single bootstrap sample

```{r}

system.time( {
  result_serial <- foreach(i = 1:1000, .combine = c) %do% {
    boot_samp <- sample(gal, n, replace = TRUE)
    med_samp <- median(boot_samp)
  }
})

```

### 1000 bootstrap samples

```{r}

system.time( {
  results_large_s <- foreach(i = 1:100, .combine = c) %do% {
  vec <- numeric(1000)
  
  for (j in 1:1000) {
    s <- sample(gal, n, replace = TRUE)
    vec[j] <- median(s)
  }
  
  }
})

```

From the results above, we see that the total run time for parallel processing is higher than for serial processing. However, with larger chunks we can see that using a parallel approach significantly decreases the processing time, thereby improving efficiency compared to serial processing.

# Question 3

```{r}

c3 <- makeCluster(4)
registerDoParallel(c3)

boot <- foreach(i = 1:10, .combine = rbind) %dopar% {

  exp_samp <- rexp(50, 1)
  boot_means <- numeric(100)
    
  for (j in 1:100) {
    b_sample <- sample(exp_samp, 50, replace = TRUE)
    boot_means[j] <- mean(b_sample) 
  }
  
  sorted <- sort(boot_means)
  lower <- sorted[ceiling(0.025*100)]
  upper <- sorted[ceiling(0.975*100)]
    
  boot_int <- quantile(boot_means, c(0.025, 0.975))
  
as.numeric(lower <= 1 && upper >= 1)

}

(coverage <- mean(boot))
  
```

```{r}
stopCluster(c3)
```

# Question 4

```{r}
#| eval: false

library(iterators)

set.seed(1234)

iter1 <- irnorm(5)
iter2 <- irnorm(5)
iter3 <- irnorm(5)

c4 <- makeCluster(3)
registerDoParallel(c4)

result <- foreach(i = iter1, j = iter2, k = iter3, .combine = rbind) %dopar% { 
  c("vector 1" = i, "vector 2" = j, "vector 3" = k) 
}

 max(result[ ,1])
 max(result[ ,2])
 max(result[ ,3])

```


```{r}
#| eval: false
stopCluster(c4)
```

# Question 5

```{r}
library(parallel)
library(foreach)
library(doParallel)

set.seed(1234)
v1 <- rnorm(5)
v2 <- rnorm(5)
v3 <- rnorm(5)

vectors <- list(v1, v2, v3)

c5 <- makeCluster(3)
registerDoParallel(c5)
```

### Time for parLapply

```{r}
#| warning: false

system.time({
  res_par <- parLapply(c5, vectors, max)
})

```

### Time for replicate

```{r}

system.time({
  result_rep <- replicate(3, max(rnorm(5)))
})

```

### Time for foreach

```{r}

system.time({
  res_foreach <- foreach(v = vectors) %dopar% {
    max(v)
  }
})

```

```{r}
stopCluster(c5)
```

From the above results, we see that the replicate function is the most efficient, with the smallest elapsed time. It is a sequential function, which usually works better for smaller data. The least efficient is parLapply, which is slightly slower than the foreach function.
